Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-03-04 cluster_paper/ms.tex
update manuscript
add venn diagram
update ms

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$175\,Myr open cluster NGC\,2516.
  %TODO: Says who? Says Piskunov+08
  The classical tidal radius of the cluster is $\approx$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form in hierarchically structured molecular clouds
\citep{shu_star_1987}.  Over the first 10\,Myr, feedback effects
including protostellar outflows, photoionization, radiation pressure,
and supernova shocks disperse the gas out of the cloud
\citep{krumholz_star_2019}.  Since only a small fraction of the cloud
mass is converted into stars ($\sim 1\%$), gas dispersal enables the
majority ($\sim$90\%) of stars in the cluster to escape the cluster's
gravitational well \citep{lada_embedded_2003}.

From 10 to 1000\,Myr, the cluster remnants that survive gas dispersal
suffer an onslaught that almost always leads to dissolution.  Within
the cluster supernovae, AGB winds, and close stellar encounters lead
to additional mass loss \citep{lamers_mass_loss_2010}.  Extrinsic to
the cluster, collisions with giant molecular clouds
\citep{spitzer_disruption_1958}, and perturbations from the galactic
tide in both the radial and vertical dimensions further promote
stellar escape \citep[{\it
e.g.},][]{fukushige_timescale_2000,bergond_gravitational_2001}.

% NOTE: words in this paragraph need work
Finding the stars that have dispersed from their natal clusters into
the galactic field is of interest for a few reasons.  One is to
understand the predominant conditions under which stars and star
clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
is to understand the Sun's birth environment \citep{adams_birth_2010}.
Did the Sun form in an open cluster, and if so, what type? Is there
any hope at identifying the stars that formed near the Sun?  Open
clusters and their dispersed remnants also provide a test case for
``chemical tagging'' \citep[{\it
e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
The converse holds too: one can use chemical tagging to test the
viability of kinematic clustering methods.

@@

The detection of dispersed stellar associations, although possible
before Gaia, has now reached a breakthrough pace \citep[{\it
e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by these studies can be summarized as follows.
On one end are coeval groups with no discernable cores ({\it e.g.},
Psc Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} Sco-Cen and Vela
OB2][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
an intermediate regime: low-density halos associated with a single
overdensity, typically an open cluster that was known before Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos could correspond to tidal tails, as have
been reported in the Hyades
\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
\citep{tang_comaber_2019}.

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.}
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Difference between different clustering methods raises the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
Our analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

As part of this broader project, we focus here on a modest question:
in just a single rich southern open cluster, is the cluster halo truly
coeval with the core?  We chose NGC\,2516 for this analysis since it
is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
to facilitate photometric rotation measurements using TESS, and some
degree of spectroscopic analysis.  Its halo is decidedly
non-spherical: \citet{kounkel_untangling_2019} reported its dimensions
as $\approx20\times10\times350$\,pc.  We want to know: is the halo
real? To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

%TODO: double check this owrding.
A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt to use ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 might be better-described as ``tidal
tails'', but the primary goal of this study however is
not to verify this claim.

@@

\subsection{Gaia Astrometry}
\label{subsec:astrometry}

@@

We selected candidate \cn\ members based on those available in the
literature.  While a number of pre-Gaia catalogs were available, the
purity and accuracy of the Gaia-derived results are the current state
of the art.  We therefore adopted what we viewed as the most
interesting clustering samples to compare: those of
\citet{cantatgaudin_gaia_2018}, \citet{kounkel_untangling_2019} and
\citet{meingast_2021}.  We refer to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of the different samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.}.  While we considering performing
our own clustering analysis using the Gaia data, such an effort would
in effect only replicate the work of these investigators.  We opt to
instead use their studies as starting points.

The Gaia clustering studies each used slightly different selection
functions, with different results.
\citetalias{cantatgaudin_gaia_2018} considered stars brighter than
$G=18$\,mag within a few degrees of the cluster center, and reported
1106 candidate members of \cn.  \citetalias{kounkel_untangling_2019}
and \citetalias{meingast_2021} considered stars up to $\approx$1\,mag
fainter, and reported 3003 and 1860 members respectively.  The
unsupervised clustering techniques that each of these studies applied
to the second Gaia Data Release are discussed
Appendix~\ref{app:clustering}, as is the overlap between their
resulting membership samples.

% %FIXME 
% \citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
% XX DEGREES} centered on the cluster's position.
% Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
% Piskunov+2008, which is 8$^\circ$ at the cluster distance of
% $\approx$400\,pc, perhaps they should have considered stars further
% out.

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study.  The \citetalias{cantatgaudin_gaia_2018} members are all within
a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members span tens of degrees.  In physical units, the physical size of
the entire structure is $\approx20\times10\times350$\,pc.
As a side note, stars in the upper right quadrant of $\mu_{\alpha'}$
{\it vs.} $\mu_\delta$ were reported almost entirely by
\citet{meingast_2021}.

In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
describe the \citetalias{cantatgaudin_gaia_2018} members as the
``core'' of the cluster, and any non-overlapping
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members as the ``halo''.  These definitions imply that there are
\ncore core members, and \nhalo halo members.  This distinction is
perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
not extend their search for members out to tens of degrees from the
cluster center.  Nonetheless, the membership catalog of
\citetalias{cantatgaudin_gaia_2018} echoes that of many previous
investigators (CITE, CITE, CITE), and is consistent with the general
{\it visual} impression that one has when looking at \cn\ in the night
sky: it appears to span $\approx$2$^\circ$, and not $\approx20^\circ$.

@@

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following ``cube'' in right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module (CITE).  We allowed the number of
stars in the comparison sample to exceed that in the cluster sample by
a factor of $\approx$5, to ensure broad sampling of stellar masses and
evolutionary states.  We also required the comparison sample to not
overlap with the cluster sample, which led to the omission of 1.1\% of
the stars drawn from the volume noted above.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars not having yet reached the
    main-sequence.
    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
    The main-sequence turn-off location and evolved stars suggests a
    cluster age slightly older than the Pleiades (dark blue line).  The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 200\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed in Appendix~\ref{app:diffred}. Based on the
available evidence, we expect that both field star contamination and
differential reddening could play a role in the larger scatter of the
halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
they are: stars past the turnoff in both the halo and core are on a
consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
the precision with which the claim can be made.  The MSTO stars are
consistent with the 178 and 316 Myr models, but are bluer than the 100
Myr model.  The RGB stars are consistent only with the 178 Myr
($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
the width of the MSTO can be attributed to binary stars, we are most
interested in the blue edge. This appears most compatible with the 178
Myr model.  These results are consistent with the MSTO age range of
165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
work for a more precise model-dependent comparison.

@@

		% \subfloat{
		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
		% }
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    % The {\it top} plot shows periods against a linear scale, while the
    % {\it bottom}  shows them against a logarithmic scale.
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
    are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

First, we collected the \nkinematic Gaia DR2 source\_ids corresponding to the
kinematically-identified core and halo members.  For each source, we first retrieved all
available CDIPS light curves, on a per-sector basis. 
This yielded
\ncdips light curves, all brighter than $Rp=16$\,mag.
The $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of NGC\,2516.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.75\textwidth]{f8a.pdf}
		}

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f8b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
    velocity space, cross-matched against the rotators, in physical
    units.}
    {\it Top:} projections in position and 2d-tangential velocity
    space. {\it Bottom:} radially binned versus 3d position and
    2d velocity difference.
    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
    velocity.
		\label{fig:physical_x_rotn}
	}
\end{figure*}

@@

\appendix
\section{Clustering method details}
\label{app:clustering}

@@

{\bf TODO: ADD A VENN DIAGRAM}

>>>>

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$175\,Myr open cluster NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form from the densest filaments of 
hierarchically structured molecular clouds
\citep{shu_star_1987}.  Over the first 10\,Myr, feedback effects
including protostellar outflows, photoionization, radiation pressure,
and supernova shocks disperse the gas out of the cloud
\citep{krumholz_star_2019}.  Since only $\sim 1\%$ of the cloud
mass is converted into stars, gas dispersal enables the
majority ($\sim$90\%) of stars in the cluster to escape the cluster's
gravitational well \citep{lada_embedded_2003}.

From 10 to 1000\,Myr, the cluster remnants that survive gas dispersal
suffer an onslaught that almost always leads to dissolution.  Inside
the cluster supernovae, AGB winds, and close stellar encounters lead
to additional mass loss \citep{lamers_mass_loss_2010}.  Extrinsic to
the cluster, collisions with giant molecular clouds
\citep{spitzer_disruption_1958}, and perturbations from the galactic
tide in both the radial and vertical dimensions further promote
stellar escape \citep[{\it
e.g.},][]{fukushige_timescale_2000,bergond_gravitational_2001}.

Finding the stars that have dispersed from their natal clusters into
the galactic field is a pressing topic for a few reasons.  One is to
understand the spatial extent and duration of star formation events
\citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
is to understand the Sun's birth environment \citep{adams_birth_2010}.
Did the Sun form in an open cluster, and if so, what type? Is there
any hope at identifying the stars that formed near the Sun?  Open
clusters and their dispersed remnants also provide a test case for
``chemical tagging'' \citep[{\it
e.g.},][]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
The converse holds too: one can use chemical tagging to test the
viability of kinematic clustering methods.

@@

The detection of dispersed stellar associations, although possible
before Gaia, has now reached a breakthrough pace \citep[{\it
e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by these studies can be summarized as follows.
On one end are coeval groups with no discernable cores ({\it e.g.},
Psc Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
an intermediate regime: low-density halos associated with a single
overdensity, typically an open cluster that was known before Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos could correspond to tidal tails, as have
been reported in the Hyades
\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
\citep{tang_comaber_2019}.

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.}
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Differences between different clustering methods raise the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
Our analysis of TESS Sectors 1-13 yielded light curves of
\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

As part of this broader project, we focus here on a modest question:
in just a single southern open cluster, is the cluster halo truly
coeval with the core?  We chose NGC\,2516 for this analysis since it
is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
to facilitate photometric rotation measurements using TESS, and some
degree of spectroscopic analysis.  Its halo is decidedly
non-spherical: \citet{kounkel_untangling_2019} reported its dimensions
as $\approx20\times10\times350$\,pc.  We want to know: is the halo
real? To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

%TODO: double check this owrding.
A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt for ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 might be best-described as ``tails'',
but the primary goal of this study is
not to verify whether these tails are tidal.

@@

%\subsection{Gaia Astrometry}
%\label{subsec:astrometry}

@@

We selected candidate \cn\ members based on those reported in the
literature.  While a number of pre-Gaia 
studies have been performed, the
purity and accuracy of the Gaia-derived results are the current state
of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.  We therefore adopted what we viewed as the most
interesting clustering samples to compare: those of
\citet{cantatgaudin_gaia_2018}, \citet{kounkel_untangling_2019} and
\citet{meingast_2021}.  We refer to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of the different samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.}.  While we considered performing
our own clustering analysis using the Gaia data, such an effort would
in effect only replicate the work of these investigators.  We opt to
instead use their studies as starting points.

The Gaia clustering studies each used different selection functions,
which yielded different results.  \citetalias{cantatgaudin_gaia_2018}
considered stars brighter than $G=18$\,mag within a few degrees of the
cluster center, and reported 1106 candidate members of \cn.
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
considered stars up to $\approx$1\,mag fainter, and reported 3003 and
1860 members respectively.  The unsupervised clustering techniques
that each of these studies applied to the second Gaia data release are
discussed Appendix~\ref{app:clustering}, as is the overlap between
their resulting membership samples.

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study in the space of positions, proper motions, and radial velocity
when available.  The \citetalias{cantatgaudin_gaia_2018} members are
all within a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members (the ``halo'') span tens of degrees. 
Inverting the parallax shows that members have been reported from
line-of-sight distances ranging from $300$ to $600\,$pc.
Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
$\mu_\delta$ projection were reported almost entirely by
\citet{meingast_2021}.

In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
describe the \citetalias{cantatgaudin_gaia_2018} members as the
``core'' of the cluster, and any non-overlapping
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members as the ``halo''.  This defintion implies that there are
\ncore\ core members, and \nhalo\ halo members.  The distinction is
perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
not extend their search for members out to tens of degrees from the
cluster center.  Nonetheless, the \citetalias{cantatgaudin_gaia_2018}
membership catalog is consistent with that of many previous
investigators \citep[{\it
e.g.},][]{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, and is
consistent with the general {\it visual} impression that one has when
looking at \cn\ in the night sky: it appears to span
$\approx$2$^\circ$, not $\approx20^\circ$.

@@

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following ``cube'' in right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample, which led to
the omission of 1.1\% of the stars drawn from the volume noted above.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars not having yet reached the
    main-sequence.
    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
     The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $Bp-Rp\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 300\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed in Appendix~\ref{app:diffred}. Based on the
available evidence, we expect that both field star contamination and
differential reddening could play a role in the larger scatter of the
halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
they are: stars past the turnoff in both the halo and core are on a
consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
the precision with which the claim can be made.  The MSTO stars are
consistent with the 178 and 316 Myr models, but are bluer than the 100
Myr model.  The RGB stars are consistent only with the 178 Myr
($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
the width of the MSTO can be attributed to binary stars, we are most
interested in the blue edge. This appears most compatible with the 178
Myr model.  These results are consistent with the MSTO age range of
165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
work for a more precise model-dependent comparison.
This implies an absolute age slightly older than the Pleiades
\citep[{\it cf.}][]{mermilliod_comparative_1981}.

@@

		% \subfloat{
		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
		% }
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
    are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

%FIXME: except actually, will need to redo with the new source list
%too .......
First, we collected the \nkinematic Gaia DR2 source\_ids corresponding
to the kinematically-identified core and halo members.  For each
source, we first retrieved all available CDIPS light curves, on a
per-sector basis.  This yielded \ncdips light curves, all brighter
than $Rp=16$\,mag.  The $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $(Bp-Rp)_0$ of 2.2, or a spectral
type of $\approx$M2V, at the distance of NGC\,2516.

@@

\begin{figure}[t]
	\begin{center}
			\includegraphics[width=0.49\textwidth]{f8a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Kinematic members of NGC\,2516, cross-matched against
  rotational members.}
    The projections are in galactic position coordinates and 2d-tangential velocity
    space.
    \label{fig:physical_x_rotn}
	}
\end{figure}

\begin{figure}[t]
	\begin{center}
			\includegraphics[width=0.49\textwidth]{f8b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
  versus 3d separation from the core and 2d velocity difference.} Bins have width
  25\,pc in position, and 2.0\,km\,s$^{-1}$ in velocity.
  {\bf Uncertainties in the fraction are taken as the quadrature sum of
  the rel unc of the numerator and denominator (FIXME)}.
  Bonafide members appear to exist hundreds of parsecs from the core, and
  up to $\approx$10\,km\,s$^{-1}$ in tangential velocity separation.
  \label{fig:hist_physical_x_rotn}
	}
\end{figure}

%\begin{figure*}[t]
%	\begin{center}
%		\leavevmode
%		\subfloat{
%			\includegraphics[width=0.75\textwidth]{f8a.pdf}
%		}
%	
%		\subfloat{
%			\includegraphics[width=0.75\textwidth]{f8b.pdf}
%		}
%	\end{center}
%	\vspace{-0.7cm}
%	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
%    velocity space, cross-matched against the rotators, in physical
%    units.}
%    {\it Top:} projections in position and 2d-tangential velocity
%    space. {\it Bottom:} radially binned versus 3d position and
%    2d velocity difference.
%    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
%    velocity.
%		\label{fig:physical_x_rotn}
%	}
%\end{figure*}

@@

\appendix
\section{Clustering methods and outcomes}
\label{app:clustering}

\begin{figure*}[h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.4\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Different clustering techniques yield different
  outcomes.} \nkinematic unique candidate cluster members found
  using three different techniques are considered in this study.
  \citetalias{cantatgaudin_gaia_2018}: \citet{cantatgaudin_gaia_2018},
  \citetalias{kounkel_untangling_2019}:
  \citet{kounkel_untangling_2019}, \citetalias{meingast_2021}:
  \citet{meingast_2021}.
  \label{fig:venn}
	}
\end{figure*}

Figure~\ref{fig:venn} is a Venn diagram showing the overlap between
the three membership catalogs concatenated in this study.  99\% of the
\citetalias{cantatgaudin_gaia_2018} sample overlaps with
\citetalias{kounkel_untangling_2019}, \citetalias{meingast_2021}, or
both.  By comparison, 36\% of the \citetalias{kounkel_untangling_2019}
sample, and 15\% of the \citetalias{meingast_2021} sample are
non-overlapping.
The data---Gaia DR2---used by all the studies was the same.
What are the differences in methodology that produce these different
outcomes?

@@
